Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms

### 2018-09-12 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: update slides

\title{Comparison of dimensionality reduction schemes for derivative-free global optimization algorithms}
\author{\textbf{Vladislav~Sovrasov}}
\institute{Lobachevsky State University of Nizhni Novgorod}
\date{}

@@

\begin{frame}
  \frametitle{Shifted and rotated evolvents}
  To create a fixed amount of preimages one can use a pre-defined set of different evolvents. These evolvents could be shifted or rotated versions of the original one. Set of shifted evolvents\footnote{Strongin, R.G., Gergel, V.P., Barkalov, K.A. Parallel methods for global optimization problem
solving (in Russian), 2009} is theoretically proven to generate at least one pair of close preimages if images are close and it perform better than the set of rotated curves.
  \begin{figure}[ht]
    %\vspace*{-0.5cm}
    \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
    \subfloat{{\includegraphics[width=.35\textwidth]{rotated.pdf} }}
  \end{figure}
%  \begin{columns}
%    \begin{column}{0.5\textwidth}
%
%    \end{column}
%    \begin{column}{0.5\textwidth}
%      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
%    \end{column}
%  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Smooth evolvent}
  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could impove convergence rate \footnote{Goryachih, A. A class of smooth modification of space-filling curves for global optimization problems, NET 2016}.
  \begin{figure}[ht]
    \includegraphics[width=0.45\textwidth]{smooth.pdf}
  \end{figure}

@@

\begin{frame}
  \frametitle{Optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x^k)\) and the function \(\varphi(y(x^k))\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1}=d(t)\in (x_{t-1}, x_t)
    \end{displaymath}
    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

@@

>>>>

\title{Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms}
\author{\textbf{Konstantin~Barkalov}, \textbf{Vladislav~Sovrasov} and \textbf{Ilya~Lebedev}}
\institute{Lobachevsky State University of Nizhni Novgorod}
\date{}

@@

\begin{frame}
  \frametitle{Shifted and rotated evolvents}
  To create a fixed amount of preimages one can use a pre-defined set of different evolvents. These evolvents could be shifted or rotated versions of the original one. Set of shifted evolvents\footnote{Strongin, R.G., Gergel, V.P., Barkalov, K.A. Parallel methods for global optimization problem
solving (in Russian), 2009} is theoretically proven to generate at least one pair of close preimages if images are close and it perform better than the set of rotated curves.
  \begin{figure}[ht]
    \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
    \subfloat{{\includegraphics[width=.35\textwidth]{rotated.pdf} }}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Smooth evolvent}
  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could improve convergence rate \footnote{Goryachih, A. A class of smooth modification of space-filling curves for global optimization problems, NET 2016}.
  \begin{figure}[ht]
    \includegraphics[width=0.45\textwidth]{smooth.pdf}
  \end{figure}

@@

\begin{frame}
  \frametitle{Basic parallel optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x^k)\) and the function \(\varphi(y(x^k))\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1}=d(t)\in (x_{t-1}, x_t)
    \end{displaymath}
    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

\begin{frame}
  \frametitle{Parallel optimization method with multiple evolvents}

@@

\begin{frame}
  \frametitle{Results of applying the parallel algorithm}

  \begin{table}
    \centering
    \caption{Averaged numbers of iterations executed by the parallel algorithm for solving the test
  optimization problems}
    \label{tab:iterations}
    \begin{tabular}{cccccccc}
      \cline{3-8}\noalign{\smallskip}
      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
  \multicolumn{2}{c}{$N=5$}   \\
      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
  \textit{Hard}  \\
      \noalign{\smallskip}\hline
      I &
      \parbox{0.25\textwidth}{
      \begin{center}
      \textbf{1 cluster node}
      \end{center}		}
        & \textit{1} & 12167 & 25635 & & 20979 & 187353  \\
      &  & \textit{32} & 328 & 1268  & &   898 & 12208 \\
      \hline \noalign{\smallskip}
  II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
    & \textit{1} & 25312 & 11103 & & 1472 & 17009 \\
  &   & \textit{32} & 64 &   913 & & 47 & 345 \\
      \noalign{\smallskip}\hline	\noalign{\smallskip}
  III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
    & \textit{1}  & 810 & 4351 & & 868 & 5697  \\
  & & \textit{32} & 34  & 112  & & 35  & 868 \\
      \noalign{\smallskip}\hline
    \end{tabular}
  \end{table}
